---
title: "Wine_data"
author: "S/18/837"
date: "2024-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(corrplot)
library(ggplot2)
library(lavaan)
library(psych)
library(data.table)
library(factoextra)
library(knitr)
library(corrr)
library(ggcorrplot)
library(gridExtra)
```
# Load the dataset
```{r}
wine_data <- read.csv("../data/winequality-red.csv", sep = ";")
head(wine_data)
```
```{r}
view(wine_data)
```
#Standarize the data
```{r}
wine_data <- apply(wine_data,2, scale)
```
```{r}
KMO(wine_data)
```
#Principle component analysis
```{r}
pca_results <- prcomp(wine_data[,C(1:11)], scale. = T)
summary(pca_results)
```
```{r}
eigenVal <- get_eigenvalue(pca_results)
eigenVal
```
# Kaiser's Criterion
```{r}
num_factors <- sum(eigenVal$eigenvalue > 1)
num_factors
```
#scree plot

```{r}
fviz_eig(pca_results,addlabels = TRUE)
```
```{r}
fa.parallel(wine_data[,C(1:11)], fm="pa", fa ="fa")
```
# Graph of variables
```{r}
fviz_pca_var(pca_results,col.var = "red")
```
## Contribution of each variable
```{r}
fviz_cos2(pca_results,choice = "var",axes = 1:3)
```
## Factor Analysis For PC method

H_0 : 5 factors are sufficient
            vs
H_1 : More factors needed
```{r}
fa_results_pc <- factanal(wine_data[,C(1:11)], factors = 5, method = "principal", rotation = "varimax")
fa_results_pc
```
#factor loadings
```{r}
loadings_pc <- fa_results_pc$loadings
loadings_pc
```

# Communality
```{r}
communalities_pc <- rowSums(loadings_pc^2)
communalities_pc
```
## Factor Analysis For ML method

H_0 : 5 factors are sufficient
            vs
H_1 : More factors needed

```{r}
fa_results_ml <- factanal(wine_data[,c(1:11)], factors = 5, method = "ml", rotation = "varimax")
fa_results_ml
```
#factor loadings
```{r}
loadings_ml <- fa_results_ml$loadings
loadings_ml
```
# Communality
```{r}
communalities_ml <- rowSums(loadings_ml^2)
communalities_ml
```
---------------------------------------------------------------------------------------------------------------------------------------------
##Confirmatory Factor Analysis (CFA) 
```{r}
cfa_model <- '
    Factor1 =~ fixed.acidity +  pH
    Factor2 =~ free.sulfur.dioxide + total.sulfur.dioxide
    Factor3 =~ volatile.acidity + citric.acid + sulphates
    Factor4 =~ chlorides + alcohol
    Factor5 =~ residual.sugar + density
    
    fixed.acidity ~~ fixed.acidity
    volatile.acidity ~~ volatile.acidity
    citric.acid ~~ citric.acid
    residual.sugar ~~ residual.sugar
    chlorides ~~ chlorides
    free.sulfur.dioxide ~~ free.sulfur.dioxide
    total.sulfur.dioxide ~~ total.sulfur.dioxide
    density ~~ density
    pH ~~ pH
    sulphates ~~ sulphates
    alcohol ~~ alcohol
'
```

#Fit the CFA model
```{r}
cfa_fit <- sem(cfa_model, data = wine_data[,c(1:11)])
```


# Summarize the model fit
```{r}
summary(cfa_fit, fit.measures = TRUE)
```



